A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-20 paper/main.tex

\documentclass[twocolumn]{aastex631}

@@

\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
\newcommand\shadetext[2][]{%
  \setbox0=\hbox{{\special{pdf:literal 7 Tr }#2}}%
  \tikz[baseline=0]\path [#1] \pgfextra{\rlap{\copy0}} (0,-\dp0) rectangle (\wd0,\ht0);% 
  }
\newcommand{\gb}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{g: #1}} }
% \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
% \newcommand{\e}[1]{\textcolor{yellow}{\textbf{: #1}} }

@@

\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxies Across Cosmic Time
}

@@

\begin{abstract}
% Statement
Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles. 
% Problem 
However, many studies continue to use equivalent separation criteria to select galaxy pairs for pair fraction, merger fraction, and merger rate studies, though it is unclear that such a selection permits an equitable comparison between galaxies of different mass, nor as a function of redshift.   
% Our solution
We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time. 
% What are the timescales? 
We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), lead to equivalent merger timescales for low- and high-mass systems at all redshifts. 
Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$) leads to timescales that differ between low- and high-mass pairs. 
% Why it matters 
Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates must be done with self-consistent separation criteria that scales as a function of mass and redshift of the system. 
As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
\end{abstract}

@@

Our sample consists of major low-mass and high-mass pairs that are isolated but physically associated.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.

\subsection{Pair orbit sample}
We begin with a subset of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
A brief version of the selection routine is described here.

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive halos (by stellar mass) from FoF groups with virial mass\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Brynorman1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm M_{G}=10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from massive nearby systems that may perturb the dynamical state of the pair.

We require that subhalos that constitute a pair meet a minimum subhalo mass criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to trace their mass history~\citep{RG2015}, and find the peak halo mass of each subhalo.

Stellar masses are then assigned to each subhalo in the FoF group using the Abundance Matching prescription of \citet{Moster2013}. 
The peak halo mass and current snapshot are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z_{\mathrm{snap}})$.

In \citet{Chamberlain2024}, the abundance matching prescription is sampled 1000 times for each subhalo to account for the spread in the abundance matching prescription. 
However, for the purposes of the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
This choice allows us to more thoroughly consider the population under study, and also permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well. % is this a good enough reason? 
% else, maybe spread of number of pairs is so small, that we expect the noise of the timescale analysis to be dominated by the snapshot spacing anyways?

@@

The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} catalogs, which contain information about every FoF group and subhalo at each snapshot. 
The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.

@@

\subsubsection{Orbits} 
We calculate orbits for all mergers and non-mergers in our base sample. 
An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or Lookback Time.

@@

\subsubsection{"First Infall"}
The orbit of a pair may be technically defined at very early times, well before the primary and secondary are physically associated. 
However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group. Specifically, we will consider only the "Post-Infall" part of each pair's orbit.

We define the snapshot of "First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter. 
The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same. 
This is because there are cases of pair interactions where the secondary will enter the FoF group of the primary, have a pericentric passage, then leave the FoF group for a few gigayears before re-entering the primary FoF group and eventually merging with the primary.

An example of the orbit of a merging pair where a "common group" criteria may differ from a "post-infall" can be seen in Figure~\ref{fig:example-orbits}, which shows the variety of orbits that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of 5 pairs that merge, where solid lines show the parts of the orbit that occur while the primary and secondary share a common FoF group, and dashed lines show the parts of the merger where the FoF group of the secondary is different than that of the primary. 
The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.

@@

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time. The vertical grey line marks $z=1.5$ at a Lookback Time of $9.5\,\Gyr$.
    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
    (Top) Orbits of pairs that merge before $z=0$. Pairs chosen at $z=1.5$ on average merge at \kc{$z=XX$/within XX Gyrs}.
    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). 
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

\subsubsection{Uniqueness of pairs}
Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.

@@

\section{Pair Sample Properties}
\subsection{Number of Unique Pairs}
We calculate the number of unique pairs at a given snapshot by selecting all unique pairs at that snapshot that have already experienced first infall, and then sum the number of pairs that have yet to merge.
Thus, the total number of unique pairs at $z=2$ includes pairs that were added to the catalog at \textit{any} snapshot, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (for example, a pair that was first selected in the snapshot-selected catalogs at $z=1$) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of unique pairs at all redshifts from $z=0-3$.

In Fig.~\ref{fig:num-pairs}, we show the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and low redshift selected pairs will on average have merger timescales longer than the remaining time to the end of the simulation.

The light green and light pink shaded regions show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met. 
These lines are identical to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}.

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/3_number_pairs.png}
    \caption{\kc{change redshift range to 0-6}, add \# pairs from pears, change dash to dots}
    \label{fig:num-pairs}
    \end{center}
\end{figure}

\subsection{Merger Fraction}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}. At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\section{Results - The Mass and Redshift Dependence of Merger Timescales}
% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift. 
\subsection{"time til merger" for selected snapshots for LvH.}
Maybe also (distribution plot 2D hist? as function of selected separation)
\subsection{Cumulative distributions}
\subsubsection{Physical}
\subsubsection{Scaled}

@@

\section{Discussion}
% \subsection{Merger Rates for Dwarf Pairs}
% % - What combination of values best reproduces the merger fraction acquired from the sims?
% % Do you need the scale of merger probability as in Ventou to get the match?
% % Vicente & Casteels 2014 comparison (pair fractions plot)
% \subsection{Pairs of Dwarfs That Do Not Merge}
% % can talk about the unmerged fraction merger rates here
% \subsection{Choosing Samples at Lower and Higher Redshift}
% % can also compare to samples to merger fractions and merger rates chosen at z=1 and z=2 (plots you have already)
% \subsection{Comparing with Merger Rates for Massive Galaxy Pairs}
% % compare to massive pairs in both obs window and merger rate
% % compare back to the assumption of isolation

@@

** 
\appendix
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{}
    % \label{fig:low-vs-high}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{}
    % \label{fig:low-vs-high}
\end{figure*}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
%     \caption{}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
% \end{figure}

>>>>

\documentclass[twocolumn,linenumbers]{aastex631}

@@

\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
\newcommand\shadetext[2][]{%
  \setbox0=\hbox{{\special{pdf:literal 7 Tr }#2}}%
  \tikz[baseline=0]\path [#1] \pgfextra{\rlap{\copy0}} (0,-\dp0) rectangle (\wd0,\ht0);% 
  }
\newcommand{\ep}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{e: #1}} }
% \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
% \newcommand{\e}[1]{\textcolor{yellow}{\textbf{: #1}} }

@@

\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time
}

@@

\begin{abstract}
% Statement
Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles. 
% Problem 
Pair fraction, merger fraction, and merger rate studies of high-mass pairs have been studied extensively, both observationally and theoretically. 
However, it is unclear if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift. 
% Our solution
Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently 
%and converted from one system to another 
by scaling by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass major pairs as a function of cosmic time, and compare to the merger timescales of similarly-selected high-mass major pairs. 
Using our framework to select pairs via a scaled separation criterion leads to equivalent merger timescales for low- and high-mass systems at redshifts $z=0-6$. 
% What are the timescales? 
Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$), leads to timescales that differ by up to $\sim1\Gyr$ between low- and high-mass pairs. 
% Why it matters 
Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system. 
As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
\end{abstract}

@@

Our sample consists of major (stellar mass ratio $\M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.

\subsection{Pair orbit sample}
We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive halos (by stellar mass) from FoF groups with virial mass\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Brynorman1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm M_{G}=1\times 10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.

We require that subhalos constituting a pair meet a minimum subhalo mass criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to trace their mass history~\citep{RG2015}, and find the peak halo mass of each subhalo.

Stellar masses are then assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z)$, with $z$ the redshift of consideration.

In \citet{Chamberlain2024}, the abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass relation. 
However, for the purposes of the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
This choice permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well. % is this a good enough reason? 
Additionally, ~\cite{Chamberlain2024} found that the spread in the number of pairs selected at a given redshift is small, such that we expect the spread on the timescale analysis to be dominated by the redshift spacing of the simulation.

@@

The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each snapshot. 
The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.

@@

\subsubsection{Orbits} 
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$.
    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
    (Top) Orbits of pairs that merge before $z=0$.
    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). \kc{update fig}
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

We extract orbits for all mergers and non-mergers in our base sample. 
An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or lookback time.

@@

\subsubsection{"First Infall"}
The orbit of a pair may be technically defined at very early times, well before the primary and secondary are physically associated. 
However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group. Specifically, we will consider only the ``Post-Infall" part of each pair's orbit.

We define the snapshot of``First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter. 
The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same. 
This is because there are cases of pair interactions where the secondary will enter the FoF group of the primary, have a pericentric passage, then leave the FoF group for a few gigayears before re-entering the primary FoF group and eventually merging with the primary.

An example of the orbit of a merging pair where a "common group" criteria may differ from a "post-infall" can be seen in Figure~\ref{fig:example-orbits}, which shows the variety of orbits that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of 5 pairs that merge, where solid lines show the parts of the orbit that occur while the primary and secondary share a common FoF group, and dashed lines show the parts of the merger where the FoF group of the secondary is different than that of the primary. 
The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
The spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
% a variety of different orbits

@@

\subsubsection{Uniqueness of orbits}
Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.

@@

\section{Pair Sample Properties}
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{\kc{need caption}\kc{Change redshift limits?} The number of low-mass (green) and high-mass (pink) pairs as a function of redshift. The solid lines show the number of unique pairs at a given redshift, while dashed lines show the number of unique pairs at a given redshift that will merge before $z=0$.
    For these two lines, we have only counted pairs at a given redshift if they have not merged, but have experienced first infall at a previous time. 
    \kc{I believe} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
    %
    The thin dotted lines show the number of pairs that were selected via the pair criteria at that specific redshift. 
    }
    \label{fig:num-pairs}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
    \caption{\kc{redshift:  0-6} The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}. At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.\kc{might need to add on second set of lines, so i have both merger fraction from snapshot-selected and from orbit-selected.}}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Unique Pairs}
We calculate the number of unique pairs at a given snapshot by selecting all unique pairs at that snapshot that have already experienced first infall, and then sum the number of pairs that have yet to merge.
Thus, the total number of unique pairs at $z=2$ includes pairs that were added to the catalog at \textit{any} snapshot, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (for example, a pair that was first selected in the snapshot-selected catalogs at $z=1$) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of unique pairs at all redshifts from $z=0-3$.

Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and low redshift selected pairs will on average have merger timescales longer than the remaining time to the end of the simulation.

The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met. 
These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.

\subsection{Merger Fraction}
We calculate the merger fraction, $\rm F_{merge}$, at a given snapshot by dividing the number of unique merging pairs (dashed lines in Fig.~\ref{fig:num-pairs}) by the number of total unique pairs (solid lines). 
As before, a pair whose first infall occurs at $z=2$ and that merges at $z=1$ will be included in the merger fraction calculation for redshift $z=1-2$.

Note that the merger fraction is a measure of the fraction of isolated pairs, whose primary and secondary subhalos have shared a common FoF group, that merge before $z=0$. 
It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$, but declines to $0$ at $z=0$. 
This is due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.

% ###########################################################################################
% Results I
% ###########################################################################################

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift. 
\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{\kc{Consider what redshifts to show. Either add colorbar normalized for all panels (in counts) or add \# of halos and a percentage. Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$. 
    There is a positive correlation between separation at the selected redshift and the time until merger. 
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
    The lowest separation pairs have the shortest merger timescales.
    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift. For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-low}
    \end{center}
\end{figure*}

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{\kc{Consider what redshifts to show. Add colorbar. Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-high}
    \end{center}
\end{figure*}

\subsection{Time Until Merger vs. Separation}
% how we calculate for this section
We calculate the time until merger as a function of separation for low-mass and high-mass pairs at 6 different redshifts ($z=0.5,1,1.5,2,3,4$). 
At a given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each different pair.

% low mass example
For example, at $z=1.5$, there are XXX unique low-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$. Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.

% high mass example
Likewise, at $z=1.5$, there are XXX unique high-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$. 
Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
% might put the above in a table with more info?

% description of figures 
Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs. 
In Fig.~\ref{fig:timevssep-low}, the time until merger for low-mass pairs is positively correlated to the separation of the pair at each redshift.

The figures can be used to get approximate `time until merger' ranges for a given separation at a given redshift.

% 
\input{table-septime.tex}

\subsection{Time Until Merger vs. Redshift}

Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.

\subsubsection{Separation-selected samples}

\subsection{"time til merger" for selected snapshots for LvH.}
Maybe also (distribution plot 2D hist? as function of selected separation)

@@

\section{Discussion}

\subsection{About the number of pairs}
Can we say anything about the REASON the unique pairs decreases at low z? Does this plot mean that it's only because things merge and so we lose them? And not that it's driven by mass-accretion thing? I think so.

Need to be careful about phrasing since we didn't actually quantify the number of pairs that leave the sample due to having too large of masses at later snapshots.

\subsection{About the merger fraction}
Were low-mass pairs expected to have such a high merger rate in isolated environs? 
How do these merger fractions compare to other literature results? 
If we can make comparisons, discuss the differences. 
If we cannot make comparisons, explain why our study is done in a way that makes for comparisons unfair/undoable .

% \subsection{Merger Rates for Dwarf Pairs}
% % - What combination of values best reproduces the merger fraction acquired from the sims?
% % Do you need the scale of merger probability as in Ventou to get the match?
% % Vicente & Casteels 2014 comparison (pair fractions plot)
% \subsection{Pairs of Dwarfs That Do Not Merge}
% % can talk about the unmerged fraction merger rates here
% \subsection{Choosing Samples at Lower and Higher Redshift}
% % can also compare to samples to merger fractions and merger rates chosen at z=1 and z=2 (plots you have already)
% \subsection{Comparing with Merger Rates for Massive Galaxy Pairs}
% % compare to massive pairs in both obs window and merger rate
% % compare back to the assumption of isolation

@@

** 
\appendix
adfadfa
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
%     \caption{}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
% \end{figure}
